### 2015-07-28 paper/anglican.tex
added anglican for listings

% !TEX root =  mdpp.tex

% listings definition for Anglican programming language

\lstdefinelanguage{Anglican}%
{morekeywords={*,*1,*2,*3,*agent*,*allow-unresolved-vars*,*assert*,*clojure-version*,*command-line-args*,%
		*compile-files*,*compile-path*,*e,*err*,*file*,*flush-on-newline*,*in*,*macro-meta*,%
		*math-context*,*ns*,*out*,*print-dup*,*print-length*,*print-level*,*print-meta*,*print-readably*,%
		*read-eval*,*source-path*,*use-context-classloader*,*warn-on-reflection*,+,-,->,->>,..,/,:else,%
		<,<=,=,==,>,>=,@,accessor,aclone,add-classpath,add-watch,agent,agent-errors,aget,alength,alias,%
		all-ns,alter,alter-meta!,alter-var-root,amap,ancestors,and,apply,areduce,array-map,aset,%
		aset-boolean,aset-byte,aset-char,aset-double,aset-float,aset-int,aset-long,aset-short,assert,%
		assoc,assoc!,assoc-in,associative?,atom,await,await-for,await1,bases,bean,bigdec,bigint,binding,%
		bit-and,bit-and-not,bit-clear,bit-flip,bit-not,bit-or,bit-set,bit-shift-left,bit-shift-right,%
		bit-test,bit-xor,boolean,boolean-array,booleans,bound-fn,bound-fn*,butlast,byte,byte-array,%
		bytes,case,cast,char,char-array,char-escape-string,char-name-string,char?,chars,chunk,chunk-append,%
		chunk-buffer,chunk-cons,chunk-first,chunk-next,chunk-rest,chunked-seq?,class,class?,%
		clear-agent-errors,clojure-version,coll?,comment,commute,comp,comparator,compare,compare-and-set!,%
		compile,complement,concat,cond,condp,conj,conj!,cons,constantly,construct-proxy,contains?,count,%
		counted?,create-ns,create-struct,cycle,dec,decimal?,declare,def,defdist,definline,defm,defmacro,defmethod,%
		defmulti,defn,defn-,defonce,defquery,defprotocol,defrecord,defstruct,deftype,delay,delay?,deliver,deref,derive,%
		descendants,destructure,disj,disj!,dissoc,dissoc!,distinct,distinct?,do,do-template,doall,doc,%
		dorun,doseq,dosync,dotimes,doto,double,double-array,doubles,drop,drop-last,drop-while,empty,empty?,%
		ensure,enumeration-seq,eval,even?,every?,false,false?,ffirst,file-seq,filter,finally,find,find-doc,%
		find-ns,find-var,first,float,float-array,float?,floats,flush,fm,fn,fn?,fnext,for,force,format,future,%
		future-call,future-cancel,future-cancelled?,future-done?,future?,gen-class,gen-interface,gensym,%
		get,get-in,get-method,get-proxy-class,get-thread-bindings,get-validator,hash,hash-map,hash-set,%
		identical?,identity,if,if-let,if-not,ifn?,import,in-ns,inc,init-proxy,instance?,int,int-array,%
		integer?,interleave,intern,interpose,into,into-array,ints,io!,isa?,iterate,iterator-seq,juxt,%
		key,keys,keyword,keyword?,last,lazy-cat,lazy-seq,let,letfn,line-seq,list,list*,list?,load,load-file,%
		load-reader,load-string,loaded-libs,locking,long,long-array,longs,loop,macroexpand,macroexpand-1,%
		make-array,make-hierarchy,map,map?,mapcat,max,max-key,memfn,memoize,merge,merge-with,meta,%
		method-sig,methods,min,min-key,mod,monitor-enter,monitor-exit,name,namespace,neg?,new,newline,%
		next,nfirst,nil,nil?,nnext,not,not-any?,not-empty,not-every?,not=,ns,ns-aliases,ns-imports,%
		ns-interns,ns-map,ns-name,ns-publics,ns-refers,ns-resolve,ns-unalias,ns-unmap,nth,nthnext,num,%
		number?,observe,odd?,or,parents,partial,partition,pcalls,peek,persistent!,pmap,pop,pop!,pop-thread-bindings,%
		pos?,pr,pr-str,predict,prefer-method,prefers,primitives-classnames,print,print-ctor,print-doc,print-dup,%
		print-method,print-namespace-doc,print-simple,print-special-doc,print-str,printf,println,println-str,%
		prn,prn-str,promise,proxy,proxy-call-with-super,proxy-mappings,proxy-name,proxy-super,%
		push-thread-bindings,pvalues,query,quot,rand,rand-int,range,ratio?,rational?,rationalize,re-find,%
		re-groups,re-matcher,re-matches,re-pattern,re-seq,read,read-line,read-string,recur,reduce,ref,%
		ref-history-count,ref-max-history,ref-min-history,ref-set,refer,refer-clojure,reify,%
		release-pending-sends,rem,remove,remove-method,remove-ns,remove-watch,repeat,repeatedly,%
		replace,replicate,require,reset!,reset-meta!,resolve,rest,resultset-seq,reverse,reversible?,%
		rseq,rsubseq,sample,second,select-keys,send,send-off,seq,seq?,seque,sequence,sequential?,set,set!,%
		set-validator!,set?,short,short-array,shorts,shutdown-agents,slurp,some,sort,sort-by,sorted-map,%
		sorted-map-by,sorted-set,sorted-set-by,sorted?,special-form-anchor,special-symbol?,split-at,%
		split-with,str,stream?,string?,struct,struct-map,subs,subseq,subvec,supers,swap!,symbol,symbol?,%
		sync,syntax-symbol-anchor,take,take-last,take-nth,take-while,test,the-ns,throw,time,to-array,%
		to-array-2d,trampoline,transient,tree-seq,true,true?,try,type,unchecked-add,unchecked-dec,%
		unchecked-divide,unchecked-inc,unchecked-multiply,unchecked-negate,unchecked-remainder,%
		unchecked-subtract,underive,unquote,unquote-splicing,update-in,update-proxy,use,val,vals,%
		var,var-get,var-set,var?,vary-meta,vec,vector,vector?,when,when-first,when-let,when-not,%
		while,with-bindings,with-bindings*,with-in-str,with-loading-context,with-local-vars,%
		with-meta,with-open,with-out-str,with-precision,xml-seq,zero?,zipmap
	},%
	   sensitive,% ???
	      alsodigit=-,%
		     morecomment=[l];,%
			    morestring=[b]"%
				  }[keywords,comments,strings]%