### 2016-09-01 slides/ppdef.tex
in progress

\begin{frame}{Definition}

A \textbf{probabilistic program} is a stateful deterministic
computation $\mathcal{P}$:
\begin{itemize}
\item Initially, $\mathcal{P}$ expects no arguments.
\item On every call, $\mathcal{P}$ returns
    \begin{itemize}
        \item a distribution $F$,
        \item a distribution and a value $(G, y)$,
        \item a value $z$, 
        \item or $\bot$.
    \end{itemize}
\item Upon returning $F$, $\mathcal{P}$ expects $x\sim F$.
\item Upon returning $\bot$, $\mathcal{P}$ terminates.
\end{itemize}
A program is run by calling $\mathcal{P}$ repeatedly until termination.

The probability of each \textbf{trace} is $\propto\prod_{i=1}^{\left|\pmb{x}\right|} p_{F_i}(x_i) \prod_{j=1}^{\left|\pmb{y}\right|}p_{G_j}(y_{j})$.
\end{frame}